### 2025-06-19 manuscript/revision/response_to_editor.tex
(hopefully) final readthrough before submitting revision

\documentclass{letter}
\usepackage{xcolor}
\usepackage{hyperref}
\signature{Savita Karthikeyan}

\address{Big Data Institute\\University of Oxford\\UK}

\begin{document}

\begin{letter}{}

\opening{Dear Alison,}

We are delighted to receive such positive feedback, and thank the reviewers for their 
careful consideration of the manuscript and software, and the constructive and helpful 
input provided. Attached are the revised manuscript (\textit{paper.pdf}), 
a version of the manuscript showing the differences to the submitted version 
(\textit{diff.pdf}), and a response to the reviewers' comments 
(\textit{response\_to\_reviwers.pdf})

We have also addressed the editorial comments and ensured the following:
\begin{itemize}
    \item Code published on Zenodo and archival DOI updated 
    under Availability, in addition to code repository
    \item Manuscript adheres to the Overleaf template: \textit{OUP General Template}
    \item Supplementary material attached as a separate pdf (\textit{supplement.pdf})
    \item All relevant LaTeX files included
    \item Added Conflicts of Interest statement to the manuscript (none declared)
\end{itemize}

We look forward to hearing from you about our revised manuscript. Please let us know if you require any 
further clarification.

On behalf of all tsbrowse authors,
\closing{Sincerely,}

\end{letter}
\end{document}